\section{Conclusion}
\label{sec:conclusion} 

In this paper, we have shown that game servers can scale well with the
number of cores on a unified memory multi-processor system, even in
the case where all players must be aware of all other players and
their actions. The thread pool system balances load well between the
cores, and its queue-based nature means that no task is starved unless
the entire system lacks resources. Message passing through the
blocking queue allows objects to communicate intensively without
blocking each other. Running our prototype game, we show that the
8-core server can handle a factor of 2 more clients before the
response time becomes unacceptable.


From the research described in this paper, a series of further
experiments present themselves.
%
The relationship between linearly scaling load and quadratic load can
be tweaked in our implementation. This could answer questions about
which type of load scale better under multi-threaded implementations.
%
Another direction this work could be extended is to go beyond the
single shared memory computer used and distribute the workload
across clusters of computers. This could be achieved by implementing
cross-server communication directly in the server code, or by using
existing technology that makes cluster behave like shared memory
machines. 
%
Furthermore, all experiments described here were run with an update frequency of
10~Hz. This is good for many types of games, but different frequencies
are relevant for different games. Investigating the effects of running
with a higher or lower frequency of updates on server performance
could yield interesting results.
%
If, during the implementation of a complex game, it is shown that some
state changes must be atomic to keep the game state consistent, the
message passing nature of this implementation means that we can use
read-write-locks for any required blocking. If such cases are found
investigating how read-write-locking influence performance would be
worthwhile.
%
%Looking at the scope of garbage-collected languages, investigating how
%different garbage collectors compare for this system would be
%interesting. Investigating the newer multithreaded garbage collectors
%would be particularly interesting. Will they make the garbage
%collection overhead more manageable?  This naturally leads to the
%question of how the system will work without any garbage collection,
%by comparing the current Java implementation with a system written in
%a language with manual memory management, preferably C++.
